---
title: "Multiple Linear Regression Wizard"
# author: "Student"
# date: "Always"
output:
  flexdashboard::flex_dashboard:
    # orientation: rows
    social: menu
    vertical_layout: scroll
    source_code: embed
runtime: shiny
---
```{r global, include=FALSE}

state_data = as.data.frame(state.x77)
#Cleaning up the column names so that there are no spaces.
colnames(state_data)[4] = "Life.Exp"
colnames(state_data)[6] = "HS.Grad"

#Creating a population density variable.
state_data[,9] = round((state_data$Population*1000)/state_data$Area,2)
colnames(state_data)[9] = "Density"


library(flexdashboard)
library(shiny)
library(DT)
library(MASS)
library(caret)
library(car)
library(ggplot2)
library(plotly)
library(corrplot)

```

Model Options {.sidebar}
=================================================

### **Welcome!** 

Select your dataset then choose which variable you are trying to predict. Next feel free to choose which features to include in two separate models that you can then compare.

For now, this dashboard can only handle numeric features.   

<br>

### **Select Inputs:**

```{r}
selectInput(inputId = "topredict",
            label = "1. What would you like to predict?",
            choices = names(state_data))
```
```{r}

uiOutput("firstmodelselect")

output$firstmodelselect <- renderUI({ 
  col_choices = setdiff(names(state_data),input$topredict)
  
  selectInput(inputId = "model1parameters",
            label = "2. Choose Your First Model Parameters",
            choices = col_choices,
            selected = NULL,
            multiple = TRUE)
  
 
})
```
```{r}
uiOutput("secondmodelselect")

output$secondmodelselect <- renderUI({ 
  col_choices = setdiff(names(state_data),input$topredict)
  
  selectInput(inputId = "model2parameters",
            label = "3. Choose Your Second Model Parameters",
            choices = col_choices,
            selected = NULL,
            multiple = TRUE)
  
})


model1 = eventReactive(input$model1parameters, {
  predictors = paste(input$model1parameters, collapse=' + ')
  eq = as.formula(paste(input$topredict, ' ~ ', predictors))
  model1 = lm(formula = eq,
              data = state_data)
})

model2 = eventReactive(input$model2parameters, {
  predictors = paste(input$model2parameters, collapse=' + ')
  eq = as.formula(paste(input$topredict, ' ~ ', predictors))
  model2 = lm(formula = eq,
              data = state_data)
})

full.model = eventReactive(input$topredict, {
  eq = as.formula(paste(input$topredict, ' ~ .'))
  model = lm(formula = eq,
              data = state_data)
})

empty.model = eventReactive(input$topredict, {
  eq = as.formula(paste(input$topredict, ' ~ 1'))
  model = lm(formula = eq,
              data = state_data)
})
```

<br>

### **Congratulations!**
Now work your way through the tabs to learn more about Multiple Linear Regression!





The Data {data-orientation=rows}
=================================================

Predicted Variable
--------------------------------

```{r}

plotOutput("predictedfeaturedist")

output$predictedfeaturedist = renderPlot({
  
  ggplot(state_data, aes_string(x=as.name(input$topredict))) + 
    geom_histogram(fill='skyblue') +
    geom_freqpoly(color = 'red') + 
    theme_classic() +
    ggtitle("Distribution of predicted variable")
  
})

```

The Data
---------------------------

```{r}
DT::dataTableOutput("datatable")
output$datatable = DT::renderDataTable({
  
  dt = state_data[input$model1parameters]
  
  datatable(dt, rownames = T) %>%
    formatStyle(input$model1parameters,
                background = '#ecf2f1',
                fontWeight = 'bold',
                `font-color` = 'white',
                `border` = '2px solid #8a8a8a')
  
  }, options = list(scrollX = TRUE,
                  columnDefs = list(list(className = 'dt-center', targets = "_all")))
)
```


AV Plots
=================================================

AV Plots {.tabset .tabset-fade}
--------------------------------
### Added-Variable Plots

```{r}
renderPlot({
  avPlots(full.model())
})
```

### Manual AV Plot

```{r}
renderPlot({
  nopop_y = lm(Life.Exp ~ . - Population, data = state_data)
  nopop_X = lm(Population ~ . - Population, data = state_data)
  plot(nopop_X$residuals,nopop_y$residuals)
})

# renderPlot({
#   avPlots(model2())
# })
```


Correlation {data-orientation=rows}
=================================================

### Summary of Data

```{r}
# topredict, model1parameters, model2paramenters

#Basic numerical EDA for state_data dataset.
renderPrint({
  summary(state_data)
})
```

### Correlation Plot

```{r}
renderPlot({
  corrplot(cor(state_data))
})

#Basic graphical EDA for the state_data dataset.


```

Model Summaries & Plots 
=================================================

Model 1 
-------------------------------------
### Model 1 Summary

```{r}
renderPrint({
  summary(model1())
})
```

### Model 1 Plots 

```{r}
renderPlot({
  grid = par(mfrow = c(2,2))
  plot(model1())
  par(grid)
})
```

Model 2
---------------------------------------
### Model 2 Summary

```{r}
renderPrint({
  summary(model2())
})

```

### Model 2 Plots

```{r}
renderPlot({
  grid = par(mfrow = c(2,2))
  plot(model2())
  par(grid)
})


#influnce plot
#vif
#avplot
       
```

Model Comparison 
=================================================

Partial F-Test (ANOVA)

```{r}
renderPrint({
  anova(model1(),
      model2())
})
```

AIC Results

```{r}
renderPrint({
  AIC(model1(),
      model2())
})
```

BIC Results

```{r}
renderPrint({
  BIC(model1(),
      model2()) 
})
```

Stepwise Model Selection {data-orientation=rows}
=================================================

### Choices 

```{r}
selectInput(inputId = "penalty",
            label = "1. What penalty coefficient would you like?",
            choices = c('AIC: 2','BIC: log(n)'))
selectInput(inputId = "direction",
            label = "2. What direction would you like to move?",
            choices = c('forward', 'backward', 'both'))
selectInput(inputId = "start",
            label = '3. If you chose "both", where would you like to start?',
            choices = c('empty.model', 'full.model'))
```
<br>
<br>
<br>
<br>
<br>
<br>
<br>


### Results

```{r}
renderPrint({
  
  model.empty = lm(Life.Exp ~ 1, data = state_data) 
  model.full = lm(Life.Exp ~ ., data = state_data) 
  scope = list(lower = formula(model.empty), upper = formula(model.full))
  penalty = ifelse(input$penalty == 'AIC: 2', 2, log(8))
  
  if (input$direction == 'forward') {
    step(model.empty, scope, direction = input$direction, k = penalty)
  } else if (input$direction == 'backward') {
    step(model.full, scope, direction = input$direction, k = penalty)
  } else {
    if (input$start == 'empty.model') {
      step(model.empty, scope, direction = input$direction, k = penalty)
    } else {
      step(model.full, scope, direction = input$direction, k = penalty)
    }
  } 
  
})

```

Prediction 
=================================================

### Model 1 Prediction
```{r}
renderPrint({
  predicted = as.data.frame(model1()$fitted.values)
  colnames(predicted) <- c("Predicted")
  merge(state_data[input$topredict], predicted, by=0)
})
```

Code 
=================================================

```{r}

```